---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# olympicmarathon

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![CRAN status](https://www.r-pkg.org/badges/version/olympicmarathon)](https://CRAN.R-project.org/package=olympicmarathon)
<!-- badges: end -->

`olympicmarathon` is an R package that contains historic data about Olympic marathon competitors and functionality to work with the data. Previously, data about Olympic marathons has been rather sparse, and official data is usually published in PDF form which creates barriers to performing analysis on it. This package aims to facilitate the viewing and analysis of the data by compiling it into one place for easy access and creating functionality for painless exploration. The data is compiled from the [World Athletics Website](https://www.worldathletics.org/).

## Installation

You can install the current version of olympicmarathon from GitHub using the following code:

``` {r, eval = FALSE}
devtools::install_packages("mflesaker/olympicmarathon")
```

## Load

Load the package with the following code:

```{r}
library(olympicmarathon)
```

## Available Tables
 - `womens_2020_data`: dataset of the women who ran the 2020 Olympic marathon
 - `mens_2020_data`: dataset of the men who ran the 2020 Olympic marathon
 - `womens_2016_data`: dataset of the women who ran the 2016 Olympic marathon
 - `mens_2016_data`: dataset of the men who ran the 2016 Olympic marathon

## Example

This is a basic example which shows you how to solve a common problem:

```{r example}

## Generate the winner of the 2020 Olympic Women's Marathon

olympic_marathon[[1,"rank"]]
  
```

## Contributors
 - [Michelle Flesaker](https://github.com/mflesaker)
 - [Georgia Gans](https://github.com/georgiagans)
